Practical Machine Learning - Activity prediction 
========================================================

## 1. Introduction 

Based on data from activity trackers, the exercise paterns have to be predicted in this project. 

## 2. Data Description  
 
The data of this projects is sourced from The Pontifical Catholic University of Rio de Janeiro, Informatics Department & School of Computing and Communications, Lancaster University, UK [Details](http://groupware.les.inf.puc-rio.br/har#wle_paper_section#ixzz3KcyUsvj2)

>Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.

```{r}
# load libraries
library(plyr);library(lattice);library(caret);
library(randomForest);library(bitops);
library(RCurl); library(downloader);library(ggplot2);
set.seed(12345)
  # load data (disabled due to slow internet connection)
# urlTrain<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv";
# urlTest<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv";
# download(urlTrain,destfile="pml-training.csv", mode = "wb");
# download(urlTest,destfile="pml-testing.csv", mode = "wb");
dataTrainRaw<-read.csv(file="pml-training.csv",head=TRUE ) 
dataTestRaw<-read.csv(file="pml-testing.csv",head=TRUE ) 
```

The train-data consists of innitially `r nrow(dataTrainRaw)` individual observations. The test-data consists of `r nrow(dataTestRaw)` cases. Train-data and test-data share the following relevant features: 
- X
- user_name
- roll_belt 
- pitch_belt 
- yaw_belt 
- roll_dumbbell 
- pitch_dumbbell 
- yaw_dumbbell 
- roll_forearm 
- pitch_forearm 
- yaw_forearm 
- roll_arm 
- pitch_arm 
- yaw_arm

Train and test data have two unique features:
- classe .. train data 
- problem_id .. test data

## 3. Data Cleaning  

Renaming of user_name feature-values into integer values 
- adelmo->1 
- carlitos->2
- charles->3
- eurico->4
- jeremy->5
- pedro->6

Renaming of classe feature-values into integer values 
- A->1
- B->2
- C->3
- D->4
- E->5

```{r}
# definition of collumn names for train and test data
featureNames <- c('X','user_name','roll_belt','pitch_belt','yaw_belt','roll_dumbbell','pitch_dumbbell','yaw_dumbbell','roll_forearm','pitch_forearm','yaw_forearm','roll_arm','pitch_arm','yaw_arm')
trainNames <- c( featureNames , 'classe')
testNames <- c( featureNames , 'problem_id')

# filter all non-relevant collumns out
dataTrain <- dataTrainRaw[,trainNames]
dataTest <- dataTestRaw[,testNames] 

# change person names into integer values
dataTrain[["user_name"]] <-revalue(dataTrain[["user_name"]], c("adelmo"="1", "carlitos"="2", "charles"="3", "eurico"="4", "jeremy"="5", "pedro"="6"))
dataTest[["user_name"]] <-revalue(dataTest[["user_name"]], c("adelmo"="1", "carlitos"="2", "charles"="3", "eurico"="4", "jeremy"="5", "pedro"="6"))

# tranform dataTrain and dataTest into matrixes
dataTrain <- data.matrix(dataTrain)
dataTest <- data.matrix(dataTest)
```


## 4. Classification Model - Training  

For model training, a 3-fold cross-validation. Method: k-Nearest-Neighbors.

```{r}
# model training
dataTrain <-data.frame(dataTrain)
cvControl <-trainControl(method="cv", number=3, allowParallel=T, verbose=T)
modelFit <- train(dataTrain$classe ~., data=dataTrain, method = "knn",
                  preProcess='pca', trControl=cvControl , importance=TRUE);
```


## 5. Classification Model - Evaluation 

```{r}
# model testing with train data
dataTrain <-data.frame(dataTrain)
prediction <- round(predict(modelFit, newdata=dataTrain))

cm <- confusionMatrix(data =  prediction , reference = dataTrain$classe )
cm
```

Accuracy of >95% seems resonably good. 

```{r}
# out-of-sample error
outOfSampleError <- 1 - cm$overall[1]
outOfSampleError
```